---
title: "Comprehensive Examples"
author: "Gosuke Homma"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Comprehensive Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Introduction

This vignette provides comprehensive examples of using the `MultiCorrSurvBinary` package for various clinical trial simulation scenarios.

```{r load-packages}
library(MultiCorrSurvBinary)
library(dplyr)
library(ggplot2)
```

## Example 1: Oncology Phase II Trial

Simulate a single-arm Phase II oncology trial with correlated endpoints.

```{r phase2-trial}
# Phase II trial parameters
set.seed(2024)

# Generate single-arm data
phase2_data <- rCorrSurvBinary(
  nsim = 100,
  outcomes = c('OS', 'PFS', 'OR'),
  n = 80,
  mst.OS = 15,      # 15-month median OS
  mst.PFS = 8,      # 8-month median PFS
  p.OR = 0.35,      # 35% response rate
  rho.OS.PFS = 0.6, # Strong correlation between survivals
  rho.OS.OR = 0.4,  # Moderate correlation OS-response
  rho.PFS.OR = 0.5, # Moderate correlation PFS-response
  tau = 18,         # 18-month accrual
  seed = 123
)

# Summary statistics
phase2_summary <- phase2_data %>%
  summarise(
    n_sims = n_distinct(sim),
    median_OS = median(OS),
    median_PFS = median(PFS),
    response_rate = mean(OR),
    correlation_OS_PFS = cor(OS, PFS),
    correlation_OS_OR = cor(OS, OR),
    correlation_PFS_OR = cor(PFS, OR)
  )

print(phase2_summary)
```

## Example 2: Randomized Controlled Trial

Simulate a two-arm randomized controlled trial.

```{r rct-trial}
# Define RCT parameters
rct_params <- list(
  experimental = list(
    mst.OS = 24, mst.PFS = 12, p.OR = 0.5, n = 200,
    rho.OS.PFS = 0.7, rho.OS.OR = 0.3, rho.PFS.OR = 0.4
  ),
  control = list(
    mst.OS = 18, mst.PFS = 9, p.OR = 0.3, n = 200,
    rho.OS.PFS = 0.7, rho.OS.OR = 0.3, rho.PFS.OR = 0.4
  )
)

# Generate RCT data
rct_data <- rCorrSurvBinaryMultiArmSubgroup(
  nsim = 200,
  outcomes = c('OS', 'PFS', 'OR'),
  arm.params = rct_params,
  tau = 36,
  seed = 456
)

# Compare arms
arm_comparison <- rct_data %>%
  group_by(ARM) %>%
  summarise(
    n_patients = n(),
    median_OS = median(OS),
    median_PFS = median(PFS),
    response_rate = mean(OR),
    .groups = 'drop'
  )

print(arm_comparison)
```

## Example 3: Biomarker-Stratified Trial

Simulate a trial with biomarker-defined subgroups.

```{r biomarker-trial}
# Biomarker-stratified trial parameters
biomarker_params <- list(
  treatment = list(
    biomarker_positive = list(
      mst.OS = 30, mst.PFS = 18, p.OR = 0.7, n = 120,
      rho.OS.PFS = 0.6, rho.OS.OR = 0.4, rho.PFS.OR = 0.5
    ),
    biomarker_negative = list(
      mst.OS = 20, mst.PFS = 10, p.OR = 0.4, n = 80,
      rho.OS.PFS = 0.6, rho.OS.OR = 0.4, rho.PFS.OR = 0.5
    )
  ),
  control = list(
    biomarker_positive = list(
      mst.OS = 22, mst.PFS = 12, p.OR = 0.4, n = 120,
      rho.OS.PFS = 0.6, rho.OS.OR = 0.4, rho.PFS.OR = 0.5
    ),
    biomarker_negative = list(
      mst.OS = 18, mst.PFS = 8, p.OR = 0.25, n = 80,
      rho.OS.PFS = 0.6, rho.OS.OR = 0.4, rho.PFS.OR = 0.5
    )
  )
)

# Generate biomarker-stratified data
biomarker_data <- rCorrSurvBinaryMultiArmSubgroup(
  nsim = 100,
  outcomes = c('OS', 'PFS', 'OR'),
  arm.params = biomarker_params,
  tau = 24,
  seed = 789
)

# Analyze by biomarker status
biomarker_analysis <- biomarker_data %>%
  group_by(ARM, SUBGROUP) %>%
  summarise(
    n_patients = n(),
    median_OS = round(median(OS), 1),
    median_PFS = round(median(PFS), 1),
    response_rate = round(mean(OR), 3),
    .groups = 'drop'
  )

print(biomarker_analysis)
```

## Example 4: Event-Driven Analysis

Perform interim and final analyses based on event counts.

```{r event-driven-analysis}
# Use the RCT data for event-driven analysis
event_analysis <- AnalysisCorrSurvBinary(
  data = rct_data,
  E = c(50, 100, 150),  # Interim analyses at 50, 100, final at 150 events
  prioritize = "OS",
  subgroup.prioritize = c("entire"),
  alternative = "greater"
)

# Summarize results by analysis timepoint
event_summary <- event_analysis %>%
  group_by(analysis_event, outcome) %>%
  summarise(
    n_comparisons = n(),
    mean_analysis_time_months = round(mean(analysis_time) * 12, 1),
    significant_tests = sum(pvalue < 0.05, na.rm = TRUE),
    power = round(mean(pvalue < 0.05, na.rm = TRUE), 3),
    .groups = 'drop'
  )

print(event_summary)
```

## Example 5: Correlation Bounds Exploration

Explore feasible correlation ranges for different parameter combinations.

```{r correlation-exploration}
# Function to explore correlation bounds
explore_bounds <- function(mst_os, mst_pfs, p_or) {
  bounds <- CorrBounds(
    outcomes = c('OS', 'PFS', 'OR'),
    lambda.OS = log(2) / mst_os,
    lambda.PFS = log(2) / mst_pfs,
    p.OR = p_or,
    rho.OS.PFS = 0,
    rho.OS.OR = 0,
    rho.PFS.OR = 0
  )
  
  data.frame(
    mst_OS = mst_os,
    mst_PFS = mst_pfs,
    p_OR = p_or,
    OS_PFS_lower = bounds$bounds$OS.PFS$lower,
    OS_PFS_upper = bounds$bounds$OS.PFS$upper,
    OS_OR_lower = bounds$bounds$OS.OR$lower,
    OS_OR_upper = bounds$bounds$OS.OR$upper,
    PFS_OR_lower = bounds$bounds$PFS.OR$lower,
    PFS_OR_upper = bounds$bounds$PFS.OR$upper
  )
}

# Explore different parameter combinations
param_grid <- expand.grid(
  mst_os = c(12, 18, 24),
  mst_pfs = c(6, 9, 12),
  p_or = c(0.3, 0.5, 0.7)
)

bounds_results <- do.call(rbind, lapply(1:nrow(param_grid), function(i) {
  explore_bounds(param_grid$mst_os[i], param_grid$mst_pfs[i], param_grid$p_or[i])
}))

# Display bounds for different scenarios
print(head(bounds_results, 10))
```

## Example 6: Sample Size and Power Considerations

Demonstrate how correlation affects power in different scenarios.

```{r power-analysis}
# Function to simulate power for different correlation scenarios
simulate_power <- function(rho_os_pfs, n_per_arm = 100, nsim = 500) {
  # Generate data with specified correlation
  power_params <- list(
    treatment = list(
      mst.OS = 20, mst.PFS = 12, p.OR = 0.5, n = n_per_arm,
      rho.OS.PFS = rho_os_pfs, rho.OS.OR = 0.3, rho.PFS.OR = 0.4
    ),
    control = list(
      mst.OS = 15, mst.PFS = 8, p.OR = 0.3, n = n_per_arm,
      rho.OS.PFS = rho_os_pfs, rho.OS.OR = 0.3, rho.PFS.OR = 0.4
    )
  )
  
  sim_data <- rCorrSurvBinaryMultiArmSubgroup(
    nsim = nsim,
    outcomes = c('OS', 'PFS', 'OR'),
    arm.params = power_params,
    tau = 24,
    seed = 123
  )
  
  # Perform analysis
  analysis <- AnalysisCorrSurvBinary(
    data = sim_data,
    E = c(80),  # Single analysis at 80 events
    prioritize = "OS",
    subgroup.prioritize = c("entire"),
    alternative = "greater"
  )
  
  # Calculate power
  power_results <- analysis %>%
    group_by(outcome) %>%
    summarise(
      power = mean(pvalue < 0.05, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    mutate(correlation = rho_os_pfs)
  
  return(power_results)
}

# Compare power across different correlation levels
correlation_levels <- c(0.3, 0.5, 0.7)
power_comparison <- do.call(rbind, lapply(correlation_levels, simulate_power))

print(power_comparison)
```

## Example 7: Validation and Diagnostics

Validate simulation results and perform diagnostic checks.

```{r validation}
# Generate validation data
validation_data <- rCorrSurvBinary(
  nsim = 1000,
  outcomes = c('OS', 'PFS', 'OR'),
  n = 200,
  mst.OS = 18,
  mst.PFS = 10,
  p.OR = 0.4,
  rho.OS.PFS = 0.6,
  rho.OS.OR = 0.3,
  rho.PFS.OR = 0.4,
  tau = 24,
  seed = 2024
)

# Check marginal distributions
marginal_check <- validation_data %>%
  summarise(
    # OS exponential check
    theoretical_median_OS = 18,
    empirical_median_OS = median(OS),
    # PFS exponential check  
    theoretical_median_PFS = 10,
    empirical_median_PFS = median(PFS),
    # OR Bernoulli check
    theoretical_response_rate = 0.4,
    empirical_response_rate = mean(OR),
    # Correlation checks
    theoretical_cor_OS_PFS = 0.6,
    empirical_cor_OS_PFS = cor(OS, PFS),
    theoretical_cor_OS_OR = 0.3,
    empirical_cor_OS_OR = cor(OS, OR),
    theoretical_cor_PFS_OR = 0.4,
    empirical_cor_PFS_OR = cor(PFS, OR)
  )

print(marginal_check)

# Check OS >= PFS constraint
constraint_check <- validation_data %>%
  summarise(
    violations = sum(OS < PFS),
    total_observations = n(),
    violation_rate = mean(OS < PFS)
  )

print(constraint_check)
```

## Best Practices and Tips

### 1. Parameter Selection
- Always validate correlation bounds before simulation
- Consider biological plausibility of correlation values
- Use realistic survival medians based on disease context

### 2. Sample Size Planning
- Account for correlation when calculating power
- Consider event-driven analyses for survival endpoints
- Plan interim analyses carefully to control Type I error

### 3. Simulation Efficiency
- Start with small `nsim` for testing
- Use seeds for reproducibility
- Consider parallel processing for large simulations

### 4. Validation
- Check marginal distributions match specifications
- Verify correlation structures are preserved
- Ensure biological constraints (OS ≥ PFS) are maintained

This comprehensive set of examples demonstrates the flexibility and power of the `MultiCorrSurvBinary` package for clinical trial simulations.
