---
  title: "Introduction to MultiCorrSurvBinary"
author: "Gosuke Homma"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
  toc: true
toc_depth: 2
vignette: >
  %\VignetteIndexEntry{Introduction to MultiCorrSurvBinary}
%\VignetteEngine{knitr::rmarkdown}
%\VignetteEncoding{UTF-8}
---
  
  ```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 5
)
```

## Overview

The `MultiCorrSurvBinary` package provides comprehensive tools for generating correlated time-to-event and binary outcomes in clinical trial simulations. This package is particularly useful for:
  
  - Simulating correlated clinical trial endpoints
- Validating correlation structures using Fréchet-Hoeffding bounds
- Performing event-driven analyses with proper statistical testing
- Handling multiple treatment arms and subgroups

## Key Concepts

### Outcomes Supported

The package supports three types of clinical trial outcomes:
  
  1. **OS (Overall Survival)**: Time-to-event outcome following exponential distribution
2. **PFS (Progression-Free Survival)**: Time-to-event outcome following exponential distribution  
3. **OR (Objective Response)**: Binary outcome following Bernoulli distribution

### Correlation Structure

Correlations between outcomes are modeled using:
  
  - **Normal copula** for generating correlated random variables
- **Fréchet-Hoeffding bounds** for validating feasible correlation ranges
- **Analytical integration** for precise bound calculations

### OS ≥ PFS Constraint

When both OS and PFS are included, the biological constraint OS ≥ PFS is enforced with options to prioritize preserving either the OS or PFS marginal distribution.

## Basic Usage

### Load the Package

```{r load}
library(MultiCorrSurvBinary)
library(dplyr)
```

### Single Population Simulation

Generate correlated outcomes for a single population:
  
  ```{r single-pop}
# Generate correlated OS, PFS, and OR for single population
result_single <- rCorrSurvBinary(
  nsim = 10,
  outcomes = c('OS', 'PFS', 'OR'),
  n = 100,
  mst.OS = 12,      # 12-year median OS
  mst.PFS = 6,      # 6-year median PFS  
  p.OR = 0.4,       # 40% response rate
  rho.OS.PFS = 0.5, # Correlation between OS and PFS
  rho.OS.OR = 0.3,  # Correlation between OS and OR
  rho.PFS.OR = 0.4, # Correlation between PFS and OR
  tau = 24,         # 24-year accrual period
  seed = 123
)

# View first few rows
head(result_single)
```

### Correlation Bounds Validation

Validate that correlation parameters are within feasible bounds:
  
  ```{r bounds}
# Check correlation bounds
bounds_check <- CorrBounds(
  outcomes = c('OS', 'PFS', 'OR'),
  lambda.OS = log(2)/12,  # Rate parameter for OS
  lambda.PFS = log(2)/6,  # Rate parameter for PFS
  p.OR = 0.4,
  rho.OS.PFS = 0.5,
  rho.OS.OR = 0.3,
  rho.PFS.OR = 0.4
)

# Print validation results
if (bounds_check$valid) {
  cat("Correlation parameters are valid!\n")
} else {
  cat("Invalid correlations:\n")
  for (error in bounds_check$errors) {
    cat("-", error, "\n")
  }
}

# Show calculated bounds
for (bound_name in names(bounds_check$bounds)) {
  bound <- bounds_check$bounds[[bound_name]]
  cat(sprintf("%s bounds: [%.4f, %.4f]\n", 
              bound_name, bound$lower, bound$upper))
}
```

### Multiple Arms Simulation

Generate data for multiple treatment arms:
  
  ```{r multi-arm}
# Define parameters for two treatment arms
arm_params <- list(
  arm1 = list(
    mst.OS = 18, mst.PFS = 12, p.OR = 0.6, n = 150,
    rho.OS.PFS = 0.5, rho.OS.OR = 0.3, rho.PFS.OR = 0.4
  ),
  arm2 = list(
    mst.OS = 12, mst.PFS = 8, p.OR = 0.4, n = 150,
    rho.OS.PFS = 0.5, rho.OS.OR = 0.3, rho.PFS.OR = 0.4
  )
)

# Generate multi-arm data
result_multi <- rCorrSurvBinaryMultiArmSubgroup(
  nsim = 10,
  outcomes = c('OS', 'PFS', 'OR'),
  arm.params = arm_params,
  tau = 24,
  seed = 456
)

# Summary by arm
result_multi %>%
  group_by(ARM) %>%
  summarise(
    n_patients = n(),
    mean_OS = mean(OS, na.rm = TRUE),
    mean_PFS = mean(PFS, na.rm = TRUE),
    response_rate = mean(OR, na.rm = TRUE),
    .groups = 'drop'
  )
```

## Advanced Features

### Subgroup Analysis

The package supports complex subgroup structures:
  
  ```{r subgroups}
# Define parameters with subgroups
arm_params_subgroups <- list(
  arm1 = list(
    sub1 = list(
      mst.OS = 20, mst.PFS = 12, p.OR = 0.7, n = 100,
      rho.OS.PFS = 0.4, rho.OS.OR = 0.2, rho.PFS.OR = 0.3
    ),
    sub2 = list(
      mst.OS = 15, mst.PFS = 8, p.OR = 0.5, n = 50,
      rho.OS.PFS = 0.4, rho.OS.OR = 0.2, rho.PFS.OR = 0.3
    )
  ),
  arm2 = list(
    sub1 = list(
      mst.OS = 12, mst.PFS = 6, p.OR = 0.4, n = 100,
      rho.OS.PFS = 0.4, rho.OS.OR = 0.2, rho.PFS.OR = 0.3
    ),
    sub2 = list(
      mst.OS = 10, mst.PFS = 5, p.OR = 0.3, n = 50,
      rho.OS.PFS = 0.4, rho.OS.OR = 0.2, rho.PFS.OR = 0.3
    )
  )
)

# Generate subgroup data
result_subgroups <- rCorrSurvBinaryMultiArmSubgroup(
  nsim = 5,
  outcomes = c('OS', 'PFS', 'OR'),
  arm.params = arm_params_subgroups,
  tau = 18,
  seed = 789
)

# Summary by arm and subgroup
result_subgroups %>%
  group_by(ARM, SUBGROUP) %>%
  summarise(
    n_patients = n(),
    mean_OS = round(mean(OS, na.rm = TRUE), 2),
    mean_PFS = round(mean(PFS, na.rm = TRUE), 2),
    response_rate = round(mean(OR, na.rm = TRUE), 2),
    .groups = 'drop'
  )
```

### Event-Driven Analysis

Perform sequential analyses at predefined event numbers:
  
  ```{r event-driven}
# Perform event-driven analysis
analysis_results <- AnalysisCorrSurvBinary(
  data = result_multi,
  E = c(25, 50, 75),  # Event numbers for analysis
  prioritize = "OS",
  subgroup.prioritize = c("entire"),
  alternative = "greater"
)

# Summary of analysis results
analysis_results %>%
  group_by(analysis_event, outcome) %>%
  summarise(
    n_tests = n(),
    significant_tests = sum(pvalue < 0.05, na.rm = TRUE),
    mean_analysis_time = round(mean(analysis_time, na.rm = TRUE), 2),
    .groups = 'drop'
  )
```

## Key Functions Summary

- `rCorrSurvBinary()`: Generate correlated outcomes for single population
- `rCorrSurvBinaryMultiArmSubgroup()`: Generate correlated outcomes for multiple arms/subgroups  
- `CorrBounds()`: Validate correlation bounds
- `TestsSurvBinary()`: Perform statistical tests
- `AnalysisCorrSurvBinary()`: Event-driven analysis framework

## Best Practices

1. **Always validate correlation bounds** before running large simulations
2. **Use appropriate event numbers** relative to sample sizes in event-driven analyses
3. **Set seeds** for reproducible results
4. **Consider biological constraints** when specifying correlation parameters
5. **Test with small nsim** first before scaling up simulations

For more detailed examples and advanced usage, see the `examples` vignette.
